---
title: "Emperor Goose Decision Rules"
author: "Erik Osnas"
date: "May 10, 2016"
output: html_document
---
Here I add a "Harvest Regulation Objective" to the value/utility function.  Modified from *MSY.Rmd file."
```{r}
## Define important functions

HarvestPolicy <- function(n, plist = list()){
  hpol <- ifelse(n < plist[[1]], "Red", ifelse(n < plist[[2]],"Yellow","Green"))
  return(hpol)
}

ilogit <- function(x){exp(x)/(1+exp(x))}

utility <- function(n=NA, h=NA, policy=NULL, par=list(npar=c(1,1), hpar=c(1,1)), weight=c(0.5, 0.5), fn=c("logistic", "linear"), policy.parameters=c("Red"=1, "Yellow"=1, "Green"=1)){
  if(fn[1]=="exp"){util.n <- 1 - exp(-par$npar[1]*n)}
  if(fn[2]=="exp"){util.h <- 1 - exp(-par$hpar[1]*h)}
  if(fn[1]=="logistic"){util.n <- plogis(q=n, par$npar[1], par$npar[2])}
  if(fn[2]=="logistic"){util.h <- plogis(q=h, par$hpar[1], par$hpar[2])}
  if(fn[1]=="linear"){util.n <- ifelse(n <= par$npar[1], n/par$npar[1], 1)}
  if(fn[2]=="linear"){util.h <- ifelse(h <= par$hpar[1], h/par$hpar[1], 1)}
  if(fn[1]=="msy"){util.n <- rep(0,length(h))}
  if(fn[2]=="msy"){util.h <- h}
  if(length(policy)>0){
    if(sum(!policy%in%names(policy.parameters))==0){ util.h <- util.h*policy.parameters[policy]}
    else(stop("Names of policy parameters do not match policy"))
  }
  
  if(length(n)==1){
    util <- matrix(c(rep(util.n, length(h)), util.h),length(h),2)%*%matrix(weight,2,1)
  }
  if(length(h)==1){
    util <- matrix(c(util.n, rep(util.h, length(n))),length(n),2)%*%matrix(weight,2,1)
  }
  if(length(n)==length(h)){
    util <- matrix(c(util.n, util.h),length(h),2)%*%matrix(weight,2,1)
  }
  return(util)
}
```

Let's first look at a utility function.
Use the utility function that was elicited from Eric Tayor on 27 April 2016.  

```{r eval=TRUE}
# results from Elicitation of utiity function from Eric Tayor
# on 4-27-2016
u1 <- list(N = c(0,50,60,80,100, 150, 175, 200), 
           U = c(0,0, 0, 0.6, 0.75, 0.9, 1, 1))
u2 <- list(N = c(0,60,80,100,125,155, 200), 
           U = c(0,0.25,0.5,0.625, 0.75, 0.875, 1))
#pdf("plot1.pdf")
plot(u1$N, u1$U, pch=16, col=1)
points(u2$N, u2$U, pch=16, col=2)

u <- rbind(as.data.frame(u1),as.data.frame(u2))
u$N2 <- u$N*u$N
u$N3 <- u$N*u$N*u$N 

#fit1 <- lm(U~N+N2+N3-1, data = u)
#summary(fit1)
#n <- seq(0, 200, by=1)
#lines(n, coef(fit1)[1]*n+coef(fit1)[2]*(n^2)+coef(fit1)[3]*(n^3))
#doesn't quite fit, try logistic
library(bbmle)
nloglike <- function(m = 80, s=20, sigma=0.1){
  -sum(dnorm(U, plogis(N, location=m, scale=s), sigma, log=TRUE))
}

fit2 <- mle2(nloglike, start=list(m=80, s=20, sigma=0.1), data=u)
summary(fit2)
lines(1:200, plogis(1:200, location=coef(fit2)[1], scale=coef(fit2)[2]))
#dev.off()
```
That seems to look reasonable.  We had difficultly understanding how utility changed with harvest, and we ran out of time.  For now we will use a linear (risk-nutral) utility function for harvest where utility is linear for harvest <= 15000 and 1 otherwise.   

Putting both utility function together and weighting them equally gives:

```{r eval=TRUE}
num=seq(1,2e5, length=100)
har=seq(1,2e4,length=100)
reward=matrix(NA, length(num), length(har))
for(i in 1:length(har)){
  reward[,i] = utility(n=num, h=har[i], par=list(npar=coef(fit2)[1:2]*1000, hpar=c(15000, 1)))
}
#pdf("plot2.pdf")
image(reward, x=num, y=har)
text(x=175000, y=14000, labels="+", font = 2, cex = 2)
text(x=25000, y=1000, labels="-", font = 2, cex = 2)

plot(num, reward[,1])

plot(har, reward[100,])
#dev.off()
```

OK, now find the harvest rule that maximizes expected utility.

```{r echo = TRUE}
posterior <- TRUE
filename <- "summer_theta_logistic_output.txt"
Reps <- 1500   #1500 works well
T <- 200
a <- 0.0001   #functional response parameter for harvest, ensures that har --> 0 as N --> 0
d <- 0.99     #utility discounting, 1 - discount rate
if(posterior==FALSE){  #need to modify to match below at posterior == TRUE
#N0 <- rnorm(Reps, 155000, 25000)
#Rmax <- rlnorm(Reps, log(0.15), 0.1)
#CC <- runif(Reps, 155000, 250000)
#Theta <- 4*rbeta(Reps, 1, 1) + 1
#sigma.proc <- rnorm(Reps, 0.143, 0.02)  ##~ from Stehn and Fleischman
#cv.obs <- 0.1
}
if(posterior==TRUE){
  post <- dget(file=filename)
  dims <- dim(post$N.tot)
  spots <- sample(1:dims[1], Reps)
  N0 <- post$N.tot[spots,dims[2]]
  Rmax <- post$r.max[spots,1]
  CC <- post$CC[spots,1]
  Theta <- post$theta[spots,1]
  q <- post$q[spots,1]
  sigma.proc <- post$sigma.proc[spots,1]
  cv.obs <- rnorm(Reps, 0.07, 0.003)  #from lm(dat$SUM_TOT_SE~dat$SUM_TOT-1)
}
t.min <- 10000
t.max <- 50000
t.delta <- 1000

Har = list(Red = rlnorm(Reps, log(4000), 0.1), #increased for crippling loss ~25%
              Yellow = rlnorm(Reps, log(7000), 0.1), 
              Green = rlnorm(Reps, log(15000), 0.2))
plot(density(Har$Red), xlim=c(0,30000), col="red", main="Harvest Distribution Under Each Policy", lwd=2)
lines(density(Har$Yellow), col="orange", lwd=2)
lines(density(Har$Green), col="green", lwd=2)
## Question:  Is harvest modelled correctly?  for each policy should be a grand mean as above and then a year effect with smaller variance?
thresholds = seq(t.min, t.max, by=t.delta)
e.util = matrix(NA, nrow=length(thresholds), ncol=length(thresholds))

for(ii in 1:length(thresholds)){for(jj in ii:length(thresholds)){
  
#define harvest thresholds
H.pol = list(Red = thresholds[ii], Yellow = thresholds[jj])

Results = list(N=array(NA, dim=c(Reps,T)), 
               N.obs = array(NA, dim=c(Reps,T)), 
               HarPol = array(NA, dim=c(Reps,T)),
               Har = array(NA, dim=c(Reps,T)),
               Utility = array(NA, dim=c(Reps,T))
               )

for(i in 1:Reps){
  N <- N.obs<- numeric()
  N[1] <- N0[i]
  N.obs[1] <- 32550 #Sum_TOT from 2014 #rnorm(1,N[1], cv.obs*N[1])
  
  for(t in 2:(T+1)){
    HarPol <- HarvestPolicy(n = N.obs[t-1], plist=H.pol)
    Results$HarPol[i,t-1] <- HarPol
    Results$Har[i,t-1] <- (1-exp(-a*N[t-1]))*Har[[HarPol]][i]  #added functional response  
    E.N.pre <- N[t-1]*exp(Rmax[i]*(1 - (N[t-1]/CC[i])^Theta[i])) - Results$Har[i,t-1]  
    N[t] <- ifelse(E.N.pre <= 0,0,rlnorm(1,log(E.N.pre/CC[i]), sigma.proc[i])*CC[i])
    N.obs[t] <- ifelse(N[t]<=0,0,rnorm(1,N[t]*q[i], cv.obs[i]*N[t]*q[i]))
  }
  Results$N[i,] <- N[-1]
  Results$N.obs[i,] <- N.obs[-1]
  Results$Utility[i,] <- utility(n = Results$N[i,], h = Results$Har[i,], 
                                 par=list(npar=coef(fit2)[1:2]*1000, hpar=c(15000, 1)))
}
discount <- matrix(d^c(0:(T-1)), nrow=Reps, ncol=T, byrow=TRUE)
e.util[ii,jj] <- mean(apply(Results$Utility*discount,1,sum))
}}

smart.spot = which(e.util==max(e.util, na.rm=TRUE), arr.ind = TRUE)
harvest.rule=list(Red=thresholds[smart.spot[1]], Yellow=thresholds[smart.spot[2]])
harvest.rule
```

```{r}
#Simulate dynamics at smart threshold
H.pol = harvest.rule
#H.pol = list(Red = 1e10, Yellow = 1e10)
#H.pol = list(Red = 0, Yellow = 0)

Results = list(N=array(NA, dim=c(Reps,T)), 
               N.obs = array(NA, dim=c(Reps,T)), 
               HarPol = array(NA, dim=c(Reps,T)),
               Har = array(NA, dim=c(Reps,T)),
               Utility = array(NA, dim=c(Reps,T))
               )
for(i in 1:Reps){
  N <- N.obs<- numeric()
  N[1] <- N0[i]
  N.obs[1] <- 32550 #Sum_TOT from 2014 #rnorm(1,N[1], cv.obs*N[1])
  
  for(t in 2:(T+1)){
    HarPol <- HarvestPolicy(n = N.obs[t-1], plist=H.pol)
    Results$HarPol[i,t-1] <- HarPol
    Results$Har[i,t-1] <- (1-exp(-a*N[t-1]))*Har[[HarPol]][i]  #added functional response  
    E.N.pre <- N[t-1]*exp(Rmax[i]*(1 - (N[t-1]/CC[i])^Theta[i])) - Results$Har[i,t-1]  
    N[t] <- ifelse(E.N.pre <= 0,0,rlnorm(1,log(E.N.pre/CC[i]), sigma.proc[i])*CC[i])
    N.obs[t] <- ifelse(N[t]<=0,0,rnorm(1,N[t]*q[i], cv.obs[i]*N[t]*q[i]))
  }
  Results$N[i,] <- N[-1]
  Results$N.obs[i,] <- N.obs[-1]
  Results$Utility[i,] <-  utility(n = Results$N[i,], h = Results$Har[i,], 
                                 par=list(npar=coef(fit2)[1:2]*1000, hpar=c(15000, 1)))
}


p.regs <- as.matrix(table(Results$HarPol)/(T*Reps))
regs <- matrix(0, 3, 1)
row.names(regs) <- c("Red","Yellow","Green")

#library(scales)
#pdf("plot3.pdf")
n.plot = 100
plot(1,1, type="n",xlim=c(0, T), ylim=c(0, 1e5), xlab="Year", ylab="Obs. Population")
for(i in 1:min(100,Reps)){
  lines(1:T, Results$N.obs[i,], col=1)
}
abline(h=H.pol, col=c("Red", "Yellow"), lwd=3)
text(x=T*(1:dim(p.regs)[1])/10, y=9000, labels=row.names(p.regs))
text(x=T*(1:dim(p.regs)[1])/10, y=5000, labels=round(p.regs,2))

plot(1,1, type="n",xlim=c(0, T), ylim=c(0, 3e5), xlab="Year", ylab="True Population")
for(i in 1:min(100,Reps)){
  lines(1:T, Results$N[i,], col=1)
}
mtext(paste("Pr(extinction) = ",round(sum(Results$N[,100]<1)/Reps,2)), side=3, font=2, line=1)
#dev.off()

#pdf("plot3.2.pdf")
image(e.util, axes=FALSE, col=terrain.colors(10))
axis(1, at=seq(0,1, by=0.2), labels=round(seq(t.min, t.max, length=6),0))
axis(2, at=seq(0,1, by=0.2), labels=round(seq(t.min, t.max, length=6),0))
points(smart.spot[1,1]/length(thresholds), smart.spot[1,2]/length(thresholds), pch=16)
mtext("Closure Threshold (Red Zone)", 1, line=3, font=2)
mtext("Restriction Threshold (Yellow Zone)", 2, line=3, font=2)
mtext(paste("Expected Utility over ",T,"years, discount = ",1-d), 3, line=2, font=2)
#dev.off()

plot(1,1, type="n",xlim=c(0, T), ylim=c(0, 75000), xlab="Year", ylab="Population")
for(i in 1:min(10,Reps)){
  lines(1:T, Results$N[i,]*post$q[spots[i],1], col=1, lwd=2)
  lines(1:T, Results$N.obs[i,], col=2, lwd=2)
}

plot(1,1, type="n",xlim=c(0, T), ylim=c(0, 3e5), xlab="Year", ylab="Population")
for(i in 1:min(10,Reps)){
  lines(1:T, Results$N[i,], col=1, lwd=2)
  lines(1:T, Results$N.obs[i,]/post$q[spots[i],1], col=2, lwd=2)
}
#dev.off()
```
